---
title: "Exploratory Data Analysis of Air Quality"
author: "Balázs Nyári"
output: html_document
---

**Exercise 3**

- Produce a report in the form of an R Notebook
- The Notebook should include:
  - Code
  - Output generated
  - Formatted text to describe your working, including obesrvations about the dataset
- The report should include the following
  - General information about the dataset - examples of rows or data, details 
  about variables and no. observations
  - Descriptive statistics - measures of central tendency and variation for
  relevant variables (the sapply function could be used)
  - Data cleaning - replace missing values with the mean for the variable (see)
  section 7 of Practical 2)
  - Data visualisation - scatter plots, histograms and boxplots
  - Correlations between relevant variables
  - Boxplots by month (you will need to convert the month variable from an 
  integer variable to a categorical varaible - slide 53, week 3)
  - Density Plots by month (slide 48, week 3)

Examples of **rows**, details about **variables** and **number of observations**

**General Information about the dataset**
```{r}
df <- airquality
summary(df)
```

**Examples of rows**
```{r}
head(df)

```

**Details about variables and number of observations (153)**
```{r}
str(df)
```

**Descriptive Statistics**

- Measures of central tendency and variation for relevant variables
  - mean, median of the dataset essentially
- sapply function could be used

*Central Tendency* of **Wind Speeds** at LaGuardia Airport

- Display *Mean* and *Standard Deviation*
```{r}
# Mean
sapply(df[,-5:-6], na.rm=TRUE, mean)
# Standard Deviation
sapply(df[,-5:-6], na.rm=TRUE, sd)
```
**Data cleaning** - replacing missing values with the mean for the variable

- This only applies to *Solar.R* field & *Ozone*
```{r}
sum(is.na(df))
colSums(is.na(df))
```

```{r}
df$Solar.R[is.na(df$Solar.R)] <- mean(df$Solar.R, na.rm = TRUE)
df$Ozone[is.na(df$Ozone)] <- mean(df$Ozone, na.rm = TRUE)
head(df)
```
Display Mean and Standard Deviation after Replacing Missing Data

- (Note that the mean has not changed but the standard deviation is reduced)
```{r}
# Mean - After changes
sapply(df[,-5:-6], na.rm=TRUE, mean)
# Standard Deviation - After changes
sapply(df[,-5:-6], na.rm=TRUE, sd)
```

**Data visualisation** 

- scatter plots, histograms and boxplots

**Scatter Plots**
```{r}
# Pretty broken function
pairs(df[,-6], cex=0.7)
```

**Box Plots**
```{r}
boxplot(df$Solar.R, main="Boxplot of Solar Radiation", ylab="Radiation (ppb)")
boxplot(df$Wind, main="Boxplot of Windspeed", ylab="Windspeed (mph)")
boxplot(df$Temp, main="Boxplot of Temperature", ylab="Temperature (F)")
boxplot(df$Ozone, main="Boxplot of Ozone", ylab="Ozone level")
```

**Histogram**
```{r}
hist(df$Solar.R, main = "Histogram of Solar Radiation", xlab ="Radiation (ppb)")
hist(df$Wind, main = "Histogram of Windspeeds", xlab ="wind (mph)")
hist(df$Temp, main = "Histogram of Temperature", xlab ="Temperature (F)")
hist(df$Ozone, main = "Histogram of Ozone", xlab ="Ozone level")

```
**Correlation Plot**
```{r}
library(corrplot)
M <- cor(df[,-6])
corrplot(M, method='number')
```


**Boxplots by month** 

- Convert *Month* variable from an integer variable to a categorical variable
```{r}
df$Month = factor(df$Month)
str(df) # Factor w/5 levels created
```
```{r}
# boxplot(df$Month) - WRONG
# ******* USE GGPLOT! *********
library(ggplot2)
ggplot(df, aes(x=Month, y=Ozone)) + geom_boxplot(aes(fill=Month))
ggplot(df, aes(x=Month, y=Solar.R)) + geom_boxplot(aes(fill=Month))
ggplot(df, aes(x=Month, y=Wind)) + geom_boxplot(aes(fill=Month))
ggplot(df, aes(x=Month, y=Temp)) + geom_boxplot(aes(fill=Month))

```

**Density plots** by *Month*
```{r}
library(ggplot2)
# Correct!!!
ggplot(df,aes(x=Temp)) +
  geom_density(aes(fill = Month), alpha = 0.5)
```

**More Learning Required**

- Using *Ggplot* to plot boxplots!
- using *pairs* to plot a fuck ton of scatter graphs at once
- **DATA CLEANSING** - df$Ozone[is.na(df$Ozone)] <- mean(df$Ozone, na.rm = TRUE)
- sapply function - sapply(df[,-5:-6], na.rm=TRUE, mean)